---
title: "Introduction - basic use"
author: "Victor Perrier"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction - basic use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


This package allow you to use [billboard.js](https://naver.github.io/billboard.js/),
a re-usable easy interface JavaScript chart library, based on D3 v4+.

Supported chart types:

* line
* bar
* pie / donut
* scatter

The main function is `billboarder`, all charts begin with. You can add layer to your charts with function `bb_*`, these functions
correspond to a billboard option define in the [API docs](https://naver.github.io/billboard.js/release/latest/doc/).
There are helpers functions to quickly create a type of chart (`bb_barchart`, `bb_linechart`, `bb_piechart`, `bb_donutchart`, `bb_gauge`,
`bb_scatterplot`), they have to be called after `billboarder`.


## Bar chart

You can create a simple bar chart by passing a `data.frame` to `bb_barchart`, the first column will be used as the x-axis, and the second one as the y-axis :

```{r}
library("billboarder")

df <- as.data.frame(table(sample(letters[1:5], 50, TRUE)))
df

billboarder() %>% 
  bb_barchart(data = df)
```


If you want to create a grouped bar chart, you'll have to put your data in a "wide" format. You can use `tidyr::spread` or `data.table::dcast`.

```{r}
df <- as.data.frame(table(
  sample(letters[1:5], 50, TRUE),
  sample(LETTERS[1:5], 50, TRUE)
))
df <- reshape(data = df, idvar = "Var1", timevar = "Var2", direction = "wide")
df

billboarder() %>% 
  bb_barchart(data = df)
```


## Line chart

You can pass to the function `bb_linechart` a vector, in that case x-axis will be the index of that vector :

```{r}
billboarder() %>% 
  bb_linechart(data = sin(seq(-pi, pi, length.out = 10)))
```



You can change the type of line with argument `type`, for example an `area-step` :


```{r}
billboarder() %>% 
  bb_linechart(data = sin(seq(-pi, pi, length.out = 10)), type = "area-step")
```


If want to specify a variable to map to the x-axis, you had to pass a `data.frame` to the function :

```{r}
df <- data.frame(
  var_x = seq(-pi, pi, length.out = 10),
  sin = sin(seq(-pi, pi, length.out = 10))
)
df

billboarder() %>% 
  bb_linechart(data = df, x = "var_x")
```



If the first variable of the `data.frame` is a `Date` or a `POSIX`, it will be automatically mapped to the x-axis :

```{r}
df <- data.frame(
  date = seq.Date(from = as.Date("2017-06-12"), by = "day", length.out = 10),
  var = rnorm(10)
)
df

billboarder() %>% 
  bb_linechart(data = df)
```



## Scatter plot

For scatter plot, use a two column `data.frame` with function `bb_scatterplot`, or specify the x variable and the y variable (you can also specify a grouping variable) :


```{r}
billboarder() %>% 
  bb_scatterplot(data = iris[, 1:2])

billboarder() %>% 
  bb_scatterplot(data = iris, x = "Petal.Length", y = "Petal.Width", group = "Species")
```






## Pie chart

For pie chart, use `bb_piechart` with a two column `data.frame` :

```{r}
df <- data.frame(
  var = c("A", "B"),
  count = c(457, 987)
)

billboarder() %>% 
  bb_piechart(data = df)
```




## Gauge chart

Gauge only need one value :

```{r}
billboarder() %>% 
  bb_gaugechart(value = 64)
```




